Combined effects of protein expression variance and correlation on multicomponent systems

### 2020-03-14 Manuscript/Manuscript.Rmd
added marginal histograms to fig 2

```{r normalization, include=FALSE}
protImpute <- pepTidy %>% 
  group_by(peptide,run) %>% 
  mutate(intensity = intensity / intensity[label == "126"]) %>%
  ungroup() %>%
  filter(run != "ctrl", condition != "ctrl") %>%
  dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
  group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = mean(intensity, na.rm = T)) %>% ungroup() %>%
  group_by(Uniprot) %>%
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  group_by(cell) %>% 
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance) %>%
  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

@@

```{r summary_stats, include=FALSE}
protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance),
            sd = sd(abundance),
            mean = mean(abundance),
            cv = sd / mean)

@@

allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
  mutate(interval = ifelse(interval == levels(interval)[1],
                           "low",
                           ifelse(interval == levels(interval)[7],
                                  "high",
                                  "mid")))

saveRDS(allGroups,"allGroups.RDS")
```

@@

```{r figure_1, echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
allGroups <- readRDS("allGroups.RDS")

@@

p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval)) +
  geom_point(data = filter(allGroups, interval == "mid", n > 4),
             size = 2, alpha = 0.5, stroke = 0) +
  geom_point(data = filter(allGroups, interval != "mid", n > 4),
             size = 3, alpha = 0.75, stroke = 0) +
  geom_label_repel(data = filter(allGroups, accession %in% interesting_groups),
                  aes(label = name),
                  colour = 'black',
                  size = 1.5,
                  force = 5,
                  box.padding = 1) +
  #coord_cartesian(ylim=c(-10,10)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
  xlab("Mean Correlation") + ylab("Mean Coefficient of Variation")

p3 <- ggplot(allGroups, aes(x = medCors, fill = interval)) +
  annotate("rect", xmin = 0.15, xmax = Inf, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.5) +
  geom_density(data = filter(allGroups, interval != "mid"), alpha = 0.75, color = NA) +
  geom_density(data = filter(allGroups, interval == "mid"), alpha = 0.5, color = NA) +
  theme_minimal() +
  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
  theme(legend.position = "none")

p4 <- ggplot(allGroups, aes(x = medCV, fill = interval)) +
  geom_density(alpha = 0.75, color = NA) +
  theme_minimal() +
  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
  theme(legend.position = "none")

@@

# p6 <- p2 + plot_spacer() + plot_layout(widths = c(1,0.5))
#   
# (p1 + p3 + p4) / p6 + plot_layout(heights = c(1,5)) +
#   plot_annotation(tag_levels = "A") &
#   theme(legend.position = "none")

left <- (p1 / p3 / p4) + plot_layout(ncol = 1)

wrap_plots(left, p2) + plot_layout(widths = c(1,3)) +
  plot_annotation(tag_levels = "A")

>>>>

```{r normalization, include=FALSE}
protImpute <- pepTidy %>% 
  group_by(peptide,run) %>% 
  mutate(intensity = intensity / intensity[label == "126"]) %>%
  ungroup() %>%
  filter(run != "ctrl", condition != "ctrl") %>%
  dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
  group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = mean(intensity, na.rm = T)) %>% ungroup() %>%
  group_by(Uniprot) %>%
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  filter(n() > 15) %>% # comment out if using KNN
  group_by(cell) %>% 
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance)
  # spread(cell, abundance) %>%
  # kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

@@

```{r summary_stats, include=FALSE}
protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance, na.rm = TRUE),
            sd = sd(abundance, na.rm = TRUE),
            mean = mean(abundance, na.rm = TRUE),
            cv = sd / mean)

@@

allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
  mutate(int_range = interval,
         interval = ifelse(interval == levels(interval)[1],
                           "low",
                           ifelse(interval == levels(interval)[7],
                                  "high",
                                  "mid")))

#saveRDS(allGroups,"allGroups.RDS")
```

@@

```{r figure_1, echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
#allGroups <- readRDS("allGroups.RDS")

@@

p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval)) +
  geom_point(data = filter(allGroups, interval == "mid", n > 4),
             size = 2, alpha = 0.5, stroke = 0) +
  geom_point(data = filter(allGroups, interval != "mid", n > 4),
             size = 3, alpha = 0.75, stroke = 0) +
  geom_label_repel(data = filter(allGroups, accession %in% interesting_groups),
                  aes(label = name),
                  colour = 'black',
                  size = 1.5,
                  force = 5,
                  box.padding = 1) +
  #coord_cartesian(ylim=c(-10,10)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
  xlab("Mean Correlation") + ylab("Mean Coefficient of Variation") +
  xlim(-0.15,0.5) + 
  ylim(0.08,0.32)

p3 <- ggplot(allGroups, aes(x = medCors, fill = interval)) +
  annotate("rect", xmin = 0.15, xmax = Inf, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.5) +
  geom_density(data = filter(allGroups, interval != "mid"), alpha = 0.75, color = NA) +
  geom_density(data = filter(allGroups, interval == "mid"), alpha = 0.5, color = NA) +
  theme_minimal() +
  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  xlim(-0.15,0.5)

p4 <- ggplot(allGroups, aes(x = medCV, fill = interval)) +
  geom_density(alpha = 0.75, color = NA) +
  theme_minimal() +
  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
  theme(legend.position = "none",
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  xlim(0.08,0.32)

@@

# left <- (p1 / p3 / p4) + plot_layout(ncol = 1)
# 
# wrap_plots(left, p2) + plot_layout(widths = c(1,3)) +
#   plot_annotation(tag_levels = "A")

p3 + plot_spacer() + p2 + (p4 + coord_flip()) +
  plot_layout(widths = c(4,1),
              heights = c(1,4))
